\frb: Properties of VLA Bursts and Implications for Overall FRB Population

### 2016-11-18 repeater-vladata.tex
observing pattern figure almost done
figure includes all arecibo observing, but not all bursts

\newcommand{\rf}{\emph{realfast}}
\newcommand{\frb}{\emph{FRB 121102}}

@@

\title{\frb: Properties of VLA Bursts and Implications for Overall FRB Population}
\shorttitle{\frb Burst Properties}
\shortauthors{Law et al.}

@@

The discovery of repeating bursts from \frb has shown that at least some FRBs are not cataclysmic and opened potential for studying a homogenous sample of bursts...
Our recent coordinated campaign with the Very Large Array and Arecibo Observatory has made the first direct, arcsecond-scale localization of an FRB and unambiguously associated it with counterparts in other observations. This campaign detected nine bursts at the VLA from 2.5 to 3.5~GHz and many more at Arecibo, GBT, and Effelsberg. The coordination of these observations allow us to refine our picture of the physical processes in \frb.

@@

The connection of \frb to the overall population...

@@

This paper is part of a series that presents the first localization and host identification of an FRB (CITE, CITE, ...). \frb, also known as the ``repeating FRB'', was first detected in November 2012 by the Arecibo Observatory \citep{2014ApJ...790..101S}. In mid 2015, new Arecibo observations revealed a series of bursts at the same DM and sky position demonstrating that FRBs are capable of repetition \citep{2016Natur.531..202S}. Beginning in August of 2015, we made the first of nine detections of \frb with the Very Large Array (CITE) and localized it with a precision of 0.1\arcsec. Deep radio and optical observing shows that \frb is unambiguously associated with a persistent radio and optical source at a redshift of 0.193 (CITE).

\frb has now been localized three orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs:
\begin{itemize}
 \item are extremely luminous, 
 \item have a significant DM contribution from the IGM, and
 \item can be used to probe the IGM and their host galaxy.
\end{itemize}
The promise implied by the first reported FRB is now being realized \citep{2007Sci...318..777L}.

The cosmological distance for \frb could have wide-ranging implications for the FRB population as a whole. However, it has not been demonstrated that \frb is representative of the overall FRB population. In fact, the unambiguous repetition of its bursts is unique among all FRBs \citep{2015MNRAS.454..457P}, so it is natural to ask whether \frb is representative. An important first step is to demonstrate that the properties of \frb are consistent with the significant body of facts for the overall population \citep{2015MNRAS.451.3278M, 2016MPLA...3130013K}. The repeating nature of \frb provides us with several statistical tests we can use to test this connection.

We can also assume that \frb is representative and use it to constrain the physical processes at play in the overall FRB population. Although we now know that FRBs are luminous, it is not yet clear what process generates the radio bursts themselves \citep{2014PhRvD..89j3009K, 2014ApJ...785L..26L, 2016MNRAS.457..232C}. The simultaneous \frb observing campaign with the VLA, Arecibo, Effelsberg, GBT, and AMI gives a more complete picture of the spectral structure of FRB radio emission. FRB repetition also has strong implications for the number of FRB-generating systems in the universe \citep{2016MNRAS.458L..89C}.

Given that FRBs are now known to be useful probes of the IGM, there is strong justification for continuing searches for new detections and localizations. The relatively faint counterpart to \frb argues that direct localization of the radio burst will continue to be the best way to find optical hosts to measure distances. Our multi-telescope constraints on burst spectra, measurement of host properties, burst rate estimates, and other properties will inform new strategies for finding FRBs...

\section{Observations}
The data presented here were obtained from multiple programs and telescopes, but the central goal was to interferometrically localize \frb with the VLA. We ultimately settled on a strategy based on simultaneous observing between the VLA and Arecibo observatories with other observatories added on a best-effort basis. The hope was that Arecibo could be used to define times of bursts for which we could selectively search with the VLA. This would reduce the trials factor that limited sensitivity and potentially provide detections at multiple independent bands. Below, we summarize these observations, with a focus on those conducted simultaneous with VLA burst detections from \frb.

\begin{figure*}[htb]
\begin{center}
\includegraphics[width=2\columnwidth]{timeline}
\caption{Summary of observing coverage during the multi-telescope observing campaign. Days with an observation are marked with symbols, while detections of \frb are enclosed in a black circle. The black dashed line shows the VLA burst detections with simultaneous coverage with other telescopes.
\label{fig:multi}}
\end{center}
\end{figure*}

\subsection{VLA}
The \frb observing campaign started in late 2015 with a 10~hr campaign observed at 1.4~GHz in the compact D configuration. In April through May 2016, we conducted a 40~hr campaign at 3.0~GHz in the C and CnB configurations in coordination with Arecibo. We concluded with a new, 40~hr, coordinated campaign from August through September 2016 in the B configuration and during the move to the most extended A configuration. In this last campaign, the first 34 hours of VLA observations were made at 3.0~GHz, while the last 6 hours were observed at 6~GHz.

@@

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". When a candidate significance higher than 7.4$\sigma$, \rf generates a candidate visualization with an image and spectrum.

@@

3 observations during VLA bursts...

@@

Observations during VLA bursts:
...

@@

Moreover, there is weak evidence that the \frb burst rate changes during the last part of the campaign. We modeled the event detection probability as a Poisson probability with a rate parameter that scales linearly with time as $R=a + b*(\rm{MJD} - 57623)$. We then calculated the probability of measuring the observed event rate as $\Pi P_i(R)$. Figure \ref{fig:rate} shows the ($a$, $b$) parameter space with contours of 50, 90, and 95\% confidence contours.

>>>>

\newcommand{\rf}{\emph{realfast}}
\newcommand{\frb}{FRB 121102}

@@

\title{\frb: Properties of VLA Bursts and Implications for Overall FRB Population}
\shorttitle{\frb\ Burst Properties}
\shortauthors{Law et al.}

@@

The discovery of repeating bursts from \frb\ has shown that at least some FRBs are not cataclysmic and opened potential for studying a homogenous sample of bursts...
Our recent coordinated campaign with the Very Large Array and Arecibo Observatory has made the first direct, arcsecond-scale localization of an FRB and unambiguously associated it with counterparts in other observations. This campaign detected nine bursts at the VLA from 2.5 to 3.5~GHz and many more at Arecibo, GBT, and Effelsberg. The coordination of these observations allow us to refine our picture of the physical processes in \frb.

@@

The connection of \frb\ to the overall population...

@@

This paper is part of a series that presents the first localization and host identification of an FRB (CITE, CITE, ...). \frb, also known as the ``repeating FRB'', was first detected in November 2012 by the Arecibo Observatory \citep{2014ApJ...790..101S}. In mid 2015, new Arecibo observations revealed a series of bursts at the same DM and sky position demonstrating that FRBs are capable of repetition \citep{2016Natur.531..202S}. Beginning in August of 2015, we made the first of nine detections of \frb\ with the Very Large Array (CITE) and localized it with a precision of 0.1\arcsec. Deep radio and optical observing shows that \frb\ is unambiguously associated with a persistent radio and optical source at a redshift of 0.193 (CITE).

\frb\ has now been localized three orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs:
\begin{itemize}
 \item are extremely luminous, 
 \item have a significant DM contribution from the IGM, and
 \item can be used to probe the IGM and their host galaxy.
\end{itemize}
The promise implied by the first reported FRB is now being realized \citep{2007Sci...318..777L}.

The cosmological distance for \frb\ could have wide-ranging implications for the FRB population as a whole. However, it has not been demonstrated that \frb\ is representative of the overall FRB population. In fact, the unambiguous repetition of its bursts is unique among all FRBs \citep{2015MNRAS.454..457P}, so it is natural to ask whether \frb\ is representative. An important first step is to demonstrate that the properties of \frb\ are consistent with the significant body of facts for the overall population \citep{2015MNRAS.451.3278M, 2016MPLA...3130013K}. The repeating nature of \frb\ provides us with several statistical tests we can use to test this connection.

We can also assume that \frb\ is representative and use it to constrain the physical processes at play in the overall FRB population. Although we now know that FRBs are luminous, it is not yet clear what process generates the radio bursts themselves \citep{2014PhRvD..89j3009K, 2014ApJ...785L..26L, 2016MNRAS.457..232C}. The simultaneous \frb\ observing campaign with the VLA, Arecibo, Effelsberg, GBT, and AMI gives a more complete picture of the spectral structure of FRB radio emission. FRB repetition also has strong implications for the number of FRB-generating systems in the universe \citep{2016MNRAS.458L..89C}.

Given that FRBs are now known to be useful probes of the IGM, there is strong justification for continuing searches for new detections and localizations. The relatively faint counterpart to \frb\ argues that direct localization of the radio burst will continue to be the best way to find optical hosts to measure distances. Our multi-telescope constraints on burst spectra, measurement of host properties, burst rate estimates, and other properties will inform new strategies for finding FRBs...

\section{Observations}
\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{timeline}
\caption{Summary of observing coverage during the multi-telescope observing campaign. Uncircled symbols show days with observations and circles show detections of \frb. *Only the coincident Arecibo burst is currently circled*. The black dashed line shows the VLA burst detections with simultaneous coverage with other telescopes.
\label{fig:multi}}
\end{center}
\end{figure*}

The data presented here were obtained from multiple programs and telescopes, but the central goal was to interferometrically localize \frb\ with the VLA. We ultimately settled on a strategy based on simultaneous observing between the VLA and Arecibo observatories with other observatories added on a best-effort basis. The hope was that Arecibo could be used to define times of bursts for which we could selectively search with the VLA. This would reduce the trials factor that limited sensitivity and potentially provide detections at multiple independent bands. Below, we summarize these observations, with a focus on those conducted simultaneous with VLA burst detections from \frb.

\subsection{VLA}
The \frb\ observing campaign started in late 2015 with a 10~hr campaign observed at 1.4~GHz in the compact D configuration. In April through May 2016, we conducted a 40~hr campaign at 3.0~GHz in the C and CnB configurations in coordination with Arecibo. We concluded with a new, 40~hr, coordinated campaign from August through September 2016 in the B configuration and during the move to the most extended A configuration. In this last campaign, the first 34 hours of VLA observations were made at 3.0~GHz, while the last 6 hours were observed at 6~GHz.

@@

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf\ is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". When a candidate significance higher than 7.4$\sigma$, \rf\ generates a candidate visualization with an image and spectrum.

@@

4 observations during VLA bursts... One simultaneous detection. Lots more activity at Arecibo overall.

@@

2 observations during VLA bursts...

@@

Moreover, there is weak evidence that the \frb\ burst rate changes during the last part of the campaign. We modeled the event detection probability as a Poisson probability with a rate parameter that scales linearly with time as $R=a + b*(\rm{MJD} - 57623)$. We then calculated the probability of measuring the observed event rate as $\Pi P_i(R)$. Figure \ref{fig:rate} shows the ($a$, $b$) parameter space with contours of 50, 90, and 95\% confidence contours.
